---
title: "Assignment 1: Hansen AER 2013"
author: "Nathan Franz"
date: "2/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(stargazer)
library(sandwich)

setwd("C:/Users/Nathan/Documents/UT Courses/3. Spring 2019/Causal Inference/CausalInferenceCourseHomework/Assignment1_RDD") 

# Load Hansen's data
dwi = read.csv("./data/hansen_dwi.csv")
```

# Summary of Hansen's Project




# Replication

## Density of the Running Variable

If people were capable of manipulating their blood alcohol content (BAC) with the precision required to full the breathalyzer, the histogram of BAC would swell to the left of the legal threshold and drop to the right of it. That distortion of the distribution would occur because of the negative legal consequences for being charged with driving while intoxicated.

In the recreation of Hansen's BAC histogram below we do not see any evidence for sorting on the running variable. The characteristic density discontinuity is not present at BAC = 0.08, the black vertical line. 

```{r manipulation_histogram, echo=FALSE}
# add new vars: 
#   bac_min is the minimum of the two measured BACs (as used in the paper)
#   bac_overlimit is an indicator variable for whether BAC is over the 0.08% limit

dwi$over_limit <- as.numeric(eval(dwi$bac1 >= 0.08)) 



# recreate figure 1, histogram of bac_1

manipulation_histogram <- ggplot(data = dwi) +
  theme_clean() +
  theme(plot.background = element_blank()) +
  geom_histogram(aes(x=bac1), binwidth = 0.001) +
  geom_vline(xintercept = 0.08) +
  labs(title="BAC histogram") +
  xlab("BAC") +
  ylab("Frequency") +
  ylim(0,2000)
manipulation_histogram
```

# Covariates by the Running Variable



```{r results='asis', echo=FALSE}

# run regressions of covariates on running variable
lm_male_on_bac <- lm(male ~ I(bac1-0.08)*over_limit, data=dwi)
lm_white_on_bac <- lm(white ~ I(bac1-0.08)*over_limit, data=dwi)
lm_aged_on_bac <- lm(aged ~ I(bac1-0.08)*over_limit, data=dwi)
lm_acc_on_bac <- lm(acc ~ I(bac1-0.08)*over_limit, data=dwi)

models <- list(lm_male_on_bac, lm_white_on_bac, lm_aged_on_bac, lm_acc_on_bac)

# calculate robust standard errors for each regression

get_robust_ses <- function(fit) {
  sqrt(
    diag(
      sandwich::vcovHC(fit, type = "HC1")
    )
  )
}

# create Latex table

tbl1 <- stargazer(models, header = FALSE, style = "aer", 
          title = "Regression Discontinuity Estimates for Covariates",
          column.labels = c("Male", "White","Age", "Accident"),
          covariate.labels = c("DUI"),
          omit = c("Constant", "bac1", "bac1:over_limit"),
          se = lapply(models, get_robust_ses),
          dep.var.caption = '',
          dep.var.labels.include = FALSE)

```

As shown in the table above, only age shows a significant discontinuity at the BAC threshold. 

The figure below shows linear model fits. Panel C in the linear model plots illuminates the significant imbalance found for the linear regression of age on BAC presented in the table above. It seems to be an artifact of including high values of BAC, which biases the model; all the average BAC levels (represented by empty dots) within .02 percentage points of the legal BAC threshold lie above the linear fit.

There are several differences between these plots and the corresponding plots presented in Hansen's paper. Obviously, the replication plots use only one threshold BAC level, rather than the two shown in the original. These influence our regression results by including data from farther away from the threshold than Hansen uses.
 
Additionally, Panel B in the replication plot shows a positive slope below the 0.08 BAC threshold, unlike the negative slope in the original. Also, Panel C in the original has an age axis that ranges from 0.34 to 0.38, three orders of magnitude smaller than the ranges in the replication.

```{r figure2_rep, echo=FALSE}
# recreate figure 2, panels A-D

dwi$bac_bucket <- cut(dwi$bac1, seq(from=0, to=0.450, by=0.002), 
                      labels=(seq(from=0.001, to=0.449, by=0.002)),
                      include.lowest=TRUE)

dwi_sum <- dwi %>%
  group_by(bac_bucket) %>%
  summarize(
    male = mean(male),
    white = mean(white),
    acc = mean(acc),
    aged = mean(aged),
    over_limit = unique(over_limit)
  )

dwi_sum$bac1 <- as.numeric(as.character(dwi_sum$bac_bucket))

# panel A - accident at scene

panel_a <- ggplot(data = dwi, 
       aes(x = bac1, 
           y = acc, 
           group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel A. Accident at scene") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(0.05, 0.25))

# panel B - male

panel_b <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = male, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel B. Male") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(0.74, 0.83))

# panel C - age

panel_c <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = aged, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel C. Age") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(33, 38))

# panel D - white

panel_d <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = white, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel D. White") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim=c(0.03, 0.2), ylim=c(0.8, 0.9))

# Assemble panels into figure 2 replication
grid.arrange(panel_a, panel_b, panel_c, panel_d, nrow = 2)

```

The bias introduced by the linear model for age above is fixed by using a quadratic model, presented in Panel C of the figure below. However, the quadratic model also introduces a potentially disqualifying amount of variance. For example, Panel B shows a questionable fit below the legal BAC threshold.


```{r figure2_rep_quad, echo=FALSE}
# recreate figure 2, panels A-D, but with quadratic model

# panel A - accident at scene

panel_a_quadratic <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = acc, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', formula= y ~ poly(x,2), color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel A. Accident at scene") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(0.05, 0.25))

# panel B - male

panel_b_quadratic <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = male, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', formula= y ~ poly(x,2), color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel B. Male") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(0.74, 0.83))

# panel C - age

panel_c_quadratic <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = aged, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', formula= y ~ poly(x,2), color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel C. Age") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim = c(0.03, 0.2), ylim = c(33, 38))

# panel D - white

panel_d_quadratic <- ggplot(data = dwi, 
                  aes(x = bac1, 
                      y = white, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_sum, shape = 1) +
  geom_smooth(method = 'lm', formula= y ~ poly(x,2), color = 'black') +
  geom_vline(xintercept = 0.08) +
  #geom_vline(xintercept = 0.15) +
  labs(title="Panel D. White") +
  xlab("BAC") +
  ylab("") +
  coord_cartesian(xlim=c(0.03, 0.2), ylim=c(0.8, 0.9))

# Assemble panels into figure 2 replication, but quadratic
grid.arrange(panel_a_quadratic, panel_b_quadratic, 
             panel_c_quadratic, panel_d_quadratic, nrow = 2)
```

# Regression Discontinuity

```{r results='asis', echo=FALSE}

# Estimate equation 1 with recidivism as the outcome, 
#   using BAC data in [0.03,0.13]

# Make linear models, control for bac1 linearly, 
#                     interact bac1 with cutoff linearly, 
#                     interact bac1 with cutoff linearly and quadratically

linear_control <- lm(recidivism ~ 1 + male + white + aged + acc + 
                       bac1 + over_limit,
   data = dwi[dwi$bac1 >= 0.03 & dwi$bac1 <= 0.13,])

linear_interact <- lm(recidivism ~ 1 + male + white + aged + acc + 
                        bac1*over_limit,
   data = dwi[dwi$bac1 >= 0.03 & dwi$bac1 <= 0.13,])

quadratic_interact <- lm(recidivism ~ 1 + male + white + aged + acc + 
                            bac1*over_limit + I(bac1^2)*over_limit,
   data = dwi[dwi$bac1 >= 0.03 & dwi$bac1 <= 0.13,])

# Make a list of the models from above for table

rd_panel_a_models = list(linear_control, linear_interact, quadratic_interact)

# Create Latex table for R Markdown

stargazer(rd_panel_a_models, header = FALSE, style = "aer", 
          title = "Regression Discontinuity Estimates for the Effect of DUI on Recidivism, BAC in [0.03, 0.13]",
          column.labels = c("Linear Control", "With Interaction","With Quadratic Interaction"),
          covariate.labels = c("DUI"),
          omit = c("Constant", "male", "white", "aged", "acc", "bac1",
                   "I(bac1^2)", "bac1:over_limit", "over_limit:I(bac1^2)"),
          se = lapply(rd_panel_a_models, get_robust_ses),
          dep.var.caption = '',
          dep.var.labels.include = FALSE)

```

```{r results='asis', echo=FALSE}

# Estimate equation 1 with recidivism as the outcome, 
#   using BAC data in [0.055,0.105]

# Make linear models, control for bac1 linearly, 
#                     interact bac1 with cutoff linearly, 
#                     interact bac1 with cutoff linearly and quadratically

linear_control_narrow <- lm(recidivism ~ 1 + male + white + aged + acc + 
                       bac1 + over_limit,
                     data = dwi[dwi$bac1 >= 0.055 & dwi$bac1 <= 0.105,])

linear_interact_narrow <- lm(recidivism ~ 1 + male + white + aged + acc + 
                        bac1*over_limit,
                      data = dwi[dwi$bac1 >= 0.055 & dwi$bac1 <= 0.105,])

quadratic_interact_narrow <- lm(recidivism ~ 1 + male + white + aged + acc + 
                           bac1*over_limit + I(bac1^2)*over_limit,
                         data = dwi[dwi$bac1 >= 0.055 & dwi$bac1 <= 0.105,])

# Make a list of the models from above for table

rd_panel_b_models = list(linear_control_narrow, linear_interact_narrow, 
                         quadratic_interact_narrow)

# Create Latex table for R Markdown

stargazer(rd_panel_b_models, header = FALSE, style = "aer", 
          title = "Regression Discontinuity Estimates for the Effect of DUI on Recidivism, BAC in [0.055, 0.105]",
          column.labels = c("Linear Control", "With Interaction","With Quadratic Interaction"),
          covariate.labels = c("DUI"),
          omit = c("Constant", "male", "white", "aged", "acc", "bac1",
                   "I(bac1^2)", "bac1:over_limit", "over_limit:I(bac1^2)"),
          se = lapply(rd_panel_b_models, get_robust_ses),
          dep.var.caption = '',
          dep.var.labels.include = FALSE)

```

```{r figure3_rep_lin, echo=FALSE}
# Replicate Panel A of figure 3 from paper
dwi_rd_sum <- dwi[dwi$bac1<0.15,] %>%
  group_by(bac_bucket) %>%
  summarize(
    recidivism = mean(recidivism),
    over_limit = unique(as.numeric(eval(bac1 >= 0.08)))
  )

dwi_rd_sum$bac1 <- as.numeric(as.character(dwi_rd_sum$bac_bucket))

# RD - panel A, linear
rd_linear <- ggplot(data = dwi[dwi$bac1<0.15,], 
                  aes(x = bac1, 
                      y = recidivism, 
                      group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_rd_sum, shape = 1) +
  geom_smooth(method = 'lm', color = 'black') +
  geom_vline(xintercept = 0.08) +
  labs(title="Regression Discontinuity for All Offenders, Linear") +
  xlab("BAC") +
  ylab("Recidivism") +
  coord_cartesian(xlim = c(0.03, 0.15), ylim = c(0.08, 0.16))
rd_linear
```

```{r figure3_rep_quad, echo=FALSE}
# RD - panel A, quadratic
rd_quadratic <- ggplot(data = dwi[dwi$bac1<0.15,], 
                            aes(x = bac1, 
                                y = recidivism, 
                                group = over_limit)) +
  theme_clean() +
  theme(plot.background=element_blank()) +
  geom_point(data = dwi_rd_sum, shape = 1) +
  geom_smooth(method = 'lm', formula= y ~ poly(x,2), color = 'black') +
  geom_vline(xintercept = 0.08) +
  labs(title="Regression Discontinuity for All Offenders, Quadratic") +
  xlab("BAC") +
  ylab("Recidivism") +
  coord_cartesian(xlim = c(0.03, 0.15), ylim = c(0.08, 0.16))
rd_quadratic
```